---
title: "Overview of svalignments package"
author: "Robert Scharpf"
date: \today
output: BiocStyle::pdf_document
vignette: >
  %\VignetteIndexEntry{Overview of svcnvs package}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

# Introduction

The main use-cases of the `svalignments` package are (1) to identify rearranged
read pairs that support structural variants, (2) assess BLAT realignment of
selected reads for consistency with putative rearrangements and the whole-genome
aligner. Most use-cases require a BAM file. For the purpose of illustrating
examples in this vignette, we use BAM files in the `svbams` package.

# Extract rearranged read pairs from a BAM file for a region of interest

This section describes how to extract rearranged read pairs from a BAM file for
a specific genomic region of interest (e.g., a possible deletion). In the
following code chunk, we load the `svbams` package which contains a small
bam files for a region of chr15.

```{r svbams}
library(GenomicAlignments)
library(Rsamtools)
library(svbams)
library(svalignments)
path <- system.file("extdata", package="svbams")
```

First, we define the region of interest, loading the
`TxDb.Hsapiens.UCSC.hg19.refGene` package to extract the sequence length for
chromosome 15.

```{r region_of_interest}
library(TxDb.Hsapiens.UCSC.hg19.refGene)
region <- GRanges("chr15", IRanges(63201003, 63209243))
si <- seqinfo(TxDb.Hsapiens.UCSC.hg19.refGene)
seqinfo(region) <- si["chr15", ]
```

To be sure this region is big enough to capture any interesting nearby features,
we expand the region of interest to include 5kb on either side of the candidate
sequence junctions.

```{r expand}
region.expand <- GRanges("chr15", IRanges(start(region)-5e3, end(region)+5e3))
seqinfo(region.expand) <- seqinfo(region)
```

Next, we create a BamViews object that will contain metadata on the bamfiles.

```{r bamviews}
bampath <- list.files(path, pattern="cgov44t.bam$", full.names=TRUE)
bview <- BamViews(bamPaths=bampath)
```

The following wrapper for `ScanBamParams` provides a default instance of the
class with parameters for extracting improperly paired reads. Note that if
`which` is missing for `improperAlignmentParams`, all improperly paired reads
will be extracted from the BAM file. For large BAM files, this can be very slow.

```{r scanbamparams}
iparams <- improperAlignmentParams(which=region.expand, mapqFilter=30)
pparams <- properAlignmentParams(which=region.expand, mapqFilter=30)
```

Next, we extract a `GAlignmentsPairs` object encapsulating all of the improperly
and properly paired reads in this region.

```{r GAlignmentPairs}
irp <- getImproperAlignmentPairs(bview,
                                 iparams)
prp <- getProperAlignmentPairs(bview,
                               pparams)
```

For visualization, we melt the `GAlignmentPairs` to `GRanges` and coerce to a
data.frame. Depending on the size of the region, plotting all properly paired
reads may not be desirable.

```{r combine}
igr <- ga2gr(irp, is.improper=TRUE)
pgr <- ga2gr(prp, is.improper=FALSE)
pgr <- thinProperPairs(pgr, 10)
gr <- sortByRead1(c(igr, pgr))
df <- as(gr, "data.frame")
```

```{r plot_readpairs, fig.height=5, fig.width=7}
library(ggplot2)
library(scales)
colors <- c("#0072B2", "#009E73")
df$tagid <- as.numeric(df$tagid)
ggplot(df, aes(ymin=tagid-0.2, ymax=tagid+0.2,
               xmin=start/1e6, xmax=end/1e6, color=read,
               fill=read, group=tagid)) +
  geom_rect() +
  xlim(c(min(df$start), max(df$end))/1e6) +
  geom_line(aes(x=start/1e6, y=tagid)) +
  ylab("read pair index") +
  scale_x_continuous(breaks=pretty_breaks(5)) +
  geom_rect(data=as.data.frame(region),
            aes(xmin=start/1e6, xmax=end/1e6, ymin=-Inf, ymax=+Inf),
            fill="steelblue", color="transparent", alpha=0.2,
            inherit.aes=FALSE) +
  scale_color_manual(values=colors) +
  scale_fill_manual(values=colors) +
  xlab("Mb") +
  theme(axis.text.x=element_text(size=7),
        axis.text.y=element_blank()) +
  guides(fill=FALSE, color=FALSE)
```

# Extract barcoded rearranged read pairs from a targeted sequencing experiment 

Jacob TODO

## Assessing BLAT alignments

TODO
